---
title: "Homework 2"
author: "Mario Becerra"
output: html_document
---
#Market Basket Analysis of recipes around the world

In this work, a market basket analysis is made on a dataset containing recipes from around the world and their ingredients.

```{r setup}
library(ggplot2)
library(arules)
library(arulesViz)
library(dplyr)
```

```{r, echo=FALSE}
setwd('/home/mbc/Dropbox/ITAM_Dropbox/Métodos Analíticos/Tareas_Git/Tarea_02')
```

```{r}
load(file='./Tarea02_baskets.RData')
```

The file *Tarea02_baskets.RData* contains a long list of ingredients used in different recipes around the world. The format is:

1. **id**: The id of each recipe. Each recipe has a unique id.

2. **region**: The world region from which the recipe comes from.

3. **ing**: The ingredients for each recipe.

```{r}
head(baskets, 10)
```

This list was created from a previous file containing all the recipes in a different format.

```{r, eval=FALSE}
datos <- read.csv('./Notas LuisFe/clase_1_mbasket/datos/srep00196-s3.csv')
baskets <- lapply(seq(nrow(datos)), function(i){ 
    b <- as.vector(as.matrix(datos[i,2:31]))
    data.frame(id = i, region=datos[i,1],  ing = b[b!=""] )}) %>%
    rbind_all
```

Returning to the *baskets* object. The total number of ingredients in the dataset is `r length(unique(baskets$ing))`.

It is of interest to inspect the most common baskets for a particular region, not the whole world. To do this we write each in a separate text file so it can be further read with the *read.transactions* function from the package **arules**.

```{r, eval=FALSE}
write.table(baskets[baskets$region=='African', c(1,3)], file='./baskets_African.csv', sep=',', row.names=FALSE)

write.table(baskets[baskets$region=='LatinAmerican', c(1,3)], file='./baskets_LatinAmerican.csv', sep=',', row.names=FALSE)

write.table(baskets[baskets$region=='EastAsian',c(1,3)], file='./baskets_EastAsian.csv', sep=',', row.names=FALSE)

write.table(baskets[baskets$region=='EasternEuropean',c(1,3)], file='./baskets_EasternEuropean.csv', sep=',', row.names=FALSE)

write.table(baskets[baskets$region=='MiddleEastern',c(1,3)], file='./baskets_MiddleEastern.csv', sep=',', row.names=FALSE)

write.table(baskets[baskets$region=='NorthAmerican',c(1,3)], file='./baskets_NorthAmerican.csv', sep=',', row.names=FALSE)

write.table(baskets[baskets$region=='NorthernEuropean',c(1,3)], file='./baskets_NorthernEuropean.csv', sep=',', row.names=FALSE)

write.table(baskets[baskets$region=='SouthAsian',c(1,3)], file='./baskets_SouthAsian.csv', sep=',', row.names=FALSE)

write.table(baskets[baskets$region=='SoutheastAsian',c(1,3)], file='./baskets_SoutheastAsian.csv', sep=',', row.names=FALSE)
            
write.table(baskets[baskets$region=='SouthernEuropean',c(1,3)], file='./baskets_SouthernEuropean.csv', sep=',', row.names=FALSE)

write.table(baskets[baskets$region=='WesternEuropean',c(1,3)], file='./baskets_WesternEuropean.csv', sep=',', row.names=FALSE)
```

Here, we just focus on the Latin American recipes, so that's the file we'll load and analyze.

We remove from memory the previous object *baskets* containing the world wide baskets.

```{r}
rm(baskets)
```

We read the Latin American recipes file and find the 20 most frequent items by using the *apriori* algorithm.

```{r}
transac_LA <- read.transactions('./baskets_LatinAmerican.csv', format='single', cols=c(1,2), sep=',')
ap.1 <- apriori(transac_LA, parameter = list(minlen=1, maxlen=1, target='frequent itemsets'))
inspect(sort(ap.1, by='support'))
```

We see that the most frequent items are cayenne, onion, garlic, tomato and corn; which may not be a big surprise because these items are frequently used in Latin American cuisine.

We now find the most frequent itemsets containing two items.

```{r}
ap.2 <- apriori(transac_LA, parameter = list(minlen=2, maxlen=2, target='frequent itemsets'))
inspect(sort(ap.2, by='support')[1:30])
```

We see that a lot of those itemsets include onion, garlic and tomato; this is pretty obvious and intuitive because the most frequent items were precisely cayenne, onion, garlic, tomato and corn.

Now we find frequent itemsets containing three items.

```{r}
ap.3 <- apriori(transac_LA, parameter = list(minlen=3, maxlen=3, target='frequent itemsets'))
inspect(sort(ap.3, by='support')[1:30])
```

Again, no surprises here. The same items appear here, and most of them are spices typically used in Latin Americam cuisine.

In order to see a little deeper in these relations, we create association rules and filter those that have an itemset with a support bigger that 0.01 and confidence bigger than 0.3.

Let's recall that support of an itemset is simply the ratio of baskets in which that particular itemset appears, and the confidence of a rule ($S \to A$) is the number of baskets in which S and A appear divided by the number of baskets in which S appears.

```{r}
rules.1 <- apriori(transac_LA, 
                parameter = list(support = 0.01, confidence=0.20, target='rules'))
inspect(sort(rules.1, by='confidence')[1:20])
```

This means that if a recipe has celery oil, it may also include corn, tomato or garlic. This may not be very interesting because we are watching the same ingredients over and over; and that's because they are the most common items. A ore interesting value may be the lift of the rules.

The lift of a rule $S\to A$ is defined as $L(S\to a) = \frac{P(S\to a)}{P(a)}$. Bigger values of lift can be interpreted as a stronger dependency between $S$ and $A$. This may happen even if the itemsets aren't the most repeated in the dataset.

```{r}
inspect(sort(rules.1, by='lift')[1:20])
```

Here we find stronger dependencies, although the most common items keep appearing. Now we can say that these associations are stronger because we are ordering them by their lift.

We may also be interested in just a particular item, like cheese.

```{r}
rules.cheese <- subset(rules.1, lift>1 & lhs %pin% 'cheese' )
inspect(sort(rules.cheese, by='lift')[0:20])
```

We finally create a last set of rules to analyze them visually. We filter rules with a high support and a relatively high confidence; and save the object as a graph to visualize it in another software.

```{r}
rules.2 <- apriori(transac_LA, 
                parameter = list(support = 0.01, confidence=0.01, minlen=2,maxlen=2,
                                 target='rules', ext=TRUE))
rules.f <- subset(rules.2, lift > 3 & lhs.support > 0.01)
saveAsGraph(rules.f, file='rules.gml', format="gml", type='itemsets')
```









